---
title: "Solution to Exercises Week 7"
output: html_notebook
---
```{r}
library(tidyverse)
library(stringr)
```

## 14.4.2

### 1

> For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple ```str_detect()``` calls.

> - Find all words that start or end with "x".
> - Find all words that start with a vowel and end with a consonant.
> - Are there any words that contain at least one of each different vowel?

```{r}
df <- tibble(
  word = words,
  i = seq_along(words)
)

df %>%
  filter(str_detect(word, "(^x)|(x$)"))

df %>% 
  filter(str_detect(word, "(^x)") | str_detect(word, "(x$)"))
  
df %>%
  filter(str_detect(word, "^[aiouey].*[^aiouey]$"))

df %>%
  filter(str_detect(word, "^[aiouey]") & str_detect(word, "[^aiouey]$"))

# 3. Only solvable using multiple str_detects, the answer is no.
df %>%
  filter(str_detect(word, "a") & str_detect(word, "e") & str_detect(word, "i") & str_detect(word, "u") & str_detect(word, "o"))
```

### 2

> What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

Multiple words, see the tables resulting from the following mutation and arranging.

```{r}
counting <- df %>%
  mutate(
    vowels_count = str_count(word, "[aieuoy]"),
    chars_count = str_count(word, "."),
    vowels_prop = vowels_count / chars_count)

counting %>%
  arrange(desc(vowels_count))

counting %>%
  arrange(desc(vowels_prop))
```

## 14.4.3.1

### 2

> From the Harvard sentences data, extract:

> - The first word from each sentence.
> - All words ending in "ing".

```{r}
sentences %>%
  str_extract("^[A-Za-z]+") %>%
  head(10)

sentences %>%
  str_extract("[A-Za-z]+ing") %>%
  head(10)
```

## 14.4.4.1 

### 1

> Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}

numbers <- c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten")
pattern <- str_c(numbers, collapse = "|")
pattern <- str_c("(", pattern, ") ([a-zA-Z]+)")
sentences %>%
  str_subset(pattern) %>%
  str_match(pattern) %>%
  head(10)

```

## 14.4.5.1

### 1

> Replace all forward slashes in a string with backslashes.

```{r}
somestring <- "/hello///"
writeLines(str_replace_all(somestring, "/", "\\\\"))
```

### 2

> Implement a simple version of ```str_to_lower()``` using ```replace_all()```.

```{r}
# sentences %>%
#   str_to_lower()

sentences %>%
  str_replace_all(c('A' = 'a', 'B' = 'b', 'C' = 'c', 'T' = 't', 'I' = 'i')) %>% # etc. 
  head(10)

```


### 3

> Switch the first and last letters in words. Which of those strings are still words?

Those that have the same first as last letter, like yesterday, or window, but some other too.

```{r}
tibble(word = words) %>%
  mutate(switched = str_replace_all(word, "^(.)(.+)(.)$", "\\3\\2\\1")) %>%
  filter(switched %in% word)
```

## 14.4.6.1

### 1

> Split up a string like "apples, pears, and bananas" into individual components.

```{r}
x <- c("apples, pears, and bananas")

str_split(x, ",? ")
```

### 2 

> Why is it better to split up by ```boundary("word")``` than " "?

It leaves out non-alphanumeric characters, like a period or a comma.


## 28.2.1

### 1

> Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point() + 
  labs(
    x = "Engine size",
    y = "Miles per gallon on highway",
    color = "Car class",
    title = "Fuel efficiency",
    subtitle = "Highway fuel efficiency as a function of engine size by car class",
    caption = "Disclaimer: Not sure whether these data are true")
```
