---
title: "Answer Key Week 1"
output: html_notebook
---

## 3.2.4

### 3.2.4.1

Nothing, because there are no layers with geometric objects in the plot.

```{r}
library(tidyverse)
ggplot(data=mpg)
```

### 3.2.4.2

32 rows, 11 columns

```{r}
nrow(mtcars)
ncol(mtcars)
```

### 3.2.4.3

f = front-wheel drive, r = rear wheel drive, 4 = 4wd

```{r}
?mpg
```


### 3.2.4.4

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

### 3.2.4.5

Both are categorical variables. To make the plot useful, we want to see the frequencies of each cell, but only see whether there is data at all for that cell or not.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))
```

## 3.3.1

### 3.3.1.1

Because ```color="blue"``` is inside the ```aes()``` clause, instead of outside of it. Because of that, R is trying to map some variable that doesn't exist from the data frame on the color aesthetic. This is the correct syntax for setting an aesthetic attribute to a fixed value:

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

### 3.3.1.2

Categorical: manufacturer, model, trans, drv, fl, class. 
Continuous: displ, year, cyl, cty hwy

This can be seen from the output below based on the second header row, which displays the type of variable. All ```<chr>``` columns are character columns and are categorical. All ```<dbl>``` and ```<int>``` columns code for continuous variables (the abbreviation stand for numbers of double precision -- allowing decimal points -- and integer numbers -- allowing only numbers without decimal points).

```{r}
mpg
```

### 3.3.1.3

If you map a continuous variable to  shape, you will get this error: ```Error: A continuous variable can not be mapped to shape```.

What does work, for instance, is the following, where size and color seem to give a continuous representation of the mapped values, which is fine.

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = cty, y = hwy, color = year, size = displ))
```

### 3.3.1.4 

That's not a problem. Both aesthetic attributes will change as a function of the values of the single mapped variable. For instance:

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = year, size = year))
```

### 3.3.1.5

For shapes that have a border, like ```shape = 21```, ```stroke``` controls te width of the surrounding border. See the difference between the two plots below:

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = trans), shape = 21, stroke=1)
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = trans), shape = 21, stroke=2)
```

### 3.3.1.6

It's pretty cool, because it will use the values that are the result of the expression. So in this example, ```displ < 5``` yields ```TRUE``` for all engines smaller than 5 and ```FALSE``` for all engines size 5 or greater. The values ```TRUE``` and ```FALSE``` are then mapped onto the ```color``` aesthetic attribute. It shows how powerful and flexible these mappings can be.

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = displ < 5))
```
