---
title: "Solution Week 2"
output: html_notebook
---

## 3.5.1

### 3.5.1.1

It works, but you get a panel for every single value of the continuous variable. For example:

```{r}
library(tidyverse)
ggplot(mpg) +
  geom_point(aes(displ, hwy)) +
  facet_grid(~ cty)
```


### 3.5.1.2

There are some levels of ```drv``` that have no data for specific levels of ```cyl```. You can see it in the plot below because these cells don't have point.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```


### 3.5.1.3

They make faceted plots, but only with a single column in the case of ```drv ~ .``` and a single row in the case of ```. ~ cyl```. The dot tells R that there is not faceting variable on that side of the formula.

### 3.5.1.4

It's easier to identify the clusters in the facets. The disadvantage is that it is hard to see the overall effect of engine size (```displ```) on mileage (```hwy```). When you have a really large data set with overlapping clusters, it will be hard to see the individual clusters. That would be a reason to use faceting.

### 3.5.1.5

By default, ```facet_wrap``` optimizes the number of rows and columns for the panels based on the number of levels in the data. ```nrow``` gives you manual control over the number of rows, and ```ncol``` over the number of columns. ```facet_grid``` doesn't have these options because the rows and columns are controlled by the formula (```. ~ .```). To see other options that control the layout of individual panels, just read ```?facet_wrap```.

### 3.5.1.6

Because we usually look at plots on a screen where we have more space horizontally than vertically. But it should also depend a bit on whether you care more about comparing the plots on their x-axes or on their y-axes.

## 3.6.1

### 3.6.1.1

An area geom (also called ribbon). A line graph assumes a relationship between two variables on the x and y axis. Boxplots and histograms visualize distributions of data for a single variable.

### 3.6.1.2

We will see points displaying the relationship between engine size (x axis) and highway mileage (y axis). Additionally, smooth lines will show the same relationship, but without visualisation of standard error. We will see a line for each level of ```drv``` (four-wheel, front, and rear) and the colour of all points and lines will be a function of ```drv```.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

### 3.6.1.3

It removes the legend from the figure. If you remove the legend, you have more space for the actual figure. I don't think you used it earlier in the chapter.

### 3.6.1.4

It controls whether to show the standard error ribbon behind the smooth lines.

### 3.6.1.5

No, they will be the same. In one, the data set and mapping is defined locally for each layer, in the other, they are defined globally, and the layers will take the same information from the global default. 

### 3.6.1.6

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = F) 

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(aes(group = drv), se = F) 

ggplot(mpg, aes(displ, hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = F) 

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(se = F) 

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(aes(linetype = drv), se = F) 

ggplot(mpg, aes(displ, hwy)) +
  geom_point(color = 'white', size = 2, shape = 21, stroke = 1, aes(fill = drv))
```

OR, alternatively, using two layers of points:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(color = 'white', size = 5) +
  geom_point(aes(color = drv), size = 3)
```

### 3.8.1.1

The problem is overplotting: there are a lot of points that are right on top of each other. You can fix it by adding a jitter.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
```

### 3.8.1.2

```width``` and ```height``` (check with ```?geom_jitter```)

### 3.8.1.3
```geom_count``` plots a larger point if there are more points at the same location, maintaining the actual values the point should take on both axes, while ```geom_jitter``` places the points on nearby random locations so they are all visible.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```

### 3.8.1.4

```geom_boxplot``` by default uses dodge positioning, which you can see in the usage when running ```?geom_boxplot```. Here's an example:

```{r}
ggplot(mpg, aes(x = class, y = hwy, color = drv)) +
  geom_boxplot()
```
